AppPAL for Android

### 2015-09-04 paper.tex
edited

\begin{abstract}
  Users must judge apps by the information shown to them by the store.
  Employers rely on employees enforcing company policies correctly on devices at their workplace.
  Some users take time to pick apps with care whilst others do not.
  Users feel frustrated when they realise what data the apps have access to.
  They want greater control over what data they give away but they do not want to spend time reading permissions lists.
  We present AppPAL: a policy language to enforce policies about apps.
  AppPAL Policies use statements from third parties and delegation relationships to
  give us a rigorous and flexible framework for enforcing and comparing and app policies; and the trust relationships surrounding them.
  % \end{itemize}
\end{abstract}

@@

Finding the right apps can be tricky.
Users need to work out which apps are well written, which are not going to abuse their data, which will work in the way they want
  and to find the apps which suit how they want to use their device.
This can be difficult as it isn't obvious how apps use the data each has access to.

App stores give some information about their apps; descriptions of the app and screenshots as well as review scores.
Android apps show a list of permissions when they're first installed.
Soon new Android apps will display permissions requests when the app first tries to access sensitive data (such as contacts or location information).
Users do not all understand how permissions relate to their device~\cite{Felt:2012hm,Thompson:2013eb}.
Ultimately the decision which apps to use and which permissions to grant must be made by the user.

Not all apps are suitable.
A large amount of \ac{pup} is being propagated for Android devices~\cite{Truong:2014bi,Svajcer:2013tp}.
Employees are increasingly using their own phones for work (\ac{byod}).
An employer may wish to restrict which apps their employees can use.
The IT department may set a policy to prevent information leaks.
Some users worry apps will misuse their personal data;
  such a user avoids apps which can access their location, or address book.
They may apply their own personal security policy when downloading and running apps.

These policies can only be enforced by the users continuously making decisions guided by these policies when prompted about apps.
This is error-prone.
Mistakes can be made.
We believe this can be improved.
An alternative would be to write the policy down and make the computer enforce it.
To implement this we use a logic of authorization.
The policy is written in the logic and enforced by checking the policy is satisfied.

We present AppPAL, an authorization logic for reasoning about apps.
The language is an instantiation of SecPAL~\cite{Becker:2006vh} with constraints and predicates that allow us to decide which apps to run or install.
The language allows us to reason about apps using statements from third parties.
The implementation allows us to enforce the policies on a device.
We can express trust relationships amongst these parties; use constraints to do additional checks.
This lets us enforce deeper and more complex policies than existing tools such as Kirin~\cite{Enck:2009ko}.

@@

"alice" says App isRunnable
  if "work-policy" isMetBy(App)
  where BeforeHourOfDay("17") = true.
\end{lstlisting}
We can delegate policy specification to third parties or roles, and assign principals to those rolls:
\begin{lstlisting}
"alice" says "it-department" can-say 0 "work-policy" isMetBy(App).
"alice" says "bob" can-act-as "it-department".
\end{lstlisting}
We can write policies specifying which permissions an app must or must not have by its app store categorization:
    \begin{lstlisting}
"alice" says App isRunnable
   if "permissions-policy" isMetBy(App).
"alice" says "permissions-policy" isMetBy(App)
   if App isAnApp
   where
      category(App, "Photography"),
      hasPermission(App, "LOCATION") = false,
      hasPermission(App, "CAMERA") = true.
    \end{lstlisting}

@@

\noindent
Specifically our contribution is:
\begin{itemize}
  \item
    Described a scenario where an employer has a policy they want to enforce for their employees~(\autoref{sec:problem});
    and shown how the employer's policy could be implemented using AppPAL~(\autoref{sec:idea}) and installed on Alice's phone.

  \item Implemented the AppPAL language on Android and the JVM.
    We show how the language can describe properties of Android apps and Android security poilicies (\autoref{ssec:language}) and how we can check policies (\autoref{ssec:eval}).

  \item Shown the need for policy tools by demonstrating the privacy paradox holds for user privacy policies with app installations (\autoref{sec:demonstation}).
\end{itemize}

@@

To ensure this policy is met Alice promises to follow it.
She might even sign a document promising never to break the rules within the policy.
This is error-prone though---what if she makes a mistake or misses an app that breaks her policy?
Emma's policy could be implemented using existing tools.
The enterprise tool \emph{Google's Device Policy for Android}\footnote{\url{https://play.google.com/store/apps/details?id=com.google.android.apps.enterprise.dmagent}} could configure Alice's device to disallow apps from outside the Google Play Store (and in the newest version of Android let Emma set the permissions of each app on an app by app basis~\cite{AndroidMPermission:2015uq}).
AppPAL is designed to build on these tools, but make the trust and delegation relationships explicit and formalize the decision making process.
Various tools such as AppGuard~\cite{Backes:2012vm}, Dr.~Android~\&~Mr.~Hide~\cite{Jeon:2012ki} or AppFence~\cite{Hornyack:2011wq} can control the permissions or data an app can get.
These could be used to used to ensure no location data is ever obtained.
Alternately other tools like Kirin~\cite{Enck:2009ko}, Flowdroid~\cite{Fritz:2013vi} or DroidSafe~\cite{Gordon:2015et} could check that the locations are ever leaked to the web.
Various anti-virus programs are available for Android---one could be installed on Alice's phone checking against McAfee's signatures.

Whilst we could implement Emma's policy using existing tools, it is a clumsy solution.
They are not flexible: If Emma changes her policy or Alice changes jobs she needs to recheck and then to alter and remove the software on her phone accordingly.
It isn't clear what an app must do to be run, or what checks have been done if it already running on the phone.
The relationship between Alice (the user), Emma (the policy setter) and the tools Emma trusts to implement her policy isn't immediately apparent.

What happens when Alice goes home?
Emma shouldn't be able to overly control what Alice does in her private life.
Alice might not be allowed to use location tracking apps at work, but at home she might want to (to meet friends, track jogging routes or find restaurants for example).
Some mobile OSs allow app permissions to be enabled and disable at run time.
Can we enforce different policies at different times or locations?

Our research looks at the problem of picking software.
Given there are some apps you want to install and run and others you do not want to, at least some of the time:
  how can you express your preferences in such a way that they can be enforced automatically?
How can we translate policy documents from natural language into a machine checkable form?
Furthermore how can we show the trust relationships used to make these decisions clearly and precisely?

@@

We propose a change to the ecosystem, shown in \autoref{fig:ecosystem}.
People have policies which are enforced by AppPAL on their devices.
The device can make use of vetting services which run tools to infer complex properties about apps.
Users can buy from stores which ensure the only apps they see are the apps which meet their policies.
Policies can be combined

@@

In \autoref{sec:problem} Alice and Emma's had policies they wanted to enforce but no means to do so.
Instead of using several different tools to enforce Emma's policy disjointedly, we use an authorization logic.
In \autoref{lst:corporate} we give an AppPAL policy implementing Emma's app concerns on Alice's phone.
% A pictorial representation, is given in \autoref{fig:emmas_policy}; color splits statements by each speaker.

AppPAL is an instantiation of SecPAL~\cite{Becker:2006vh} for describing app policies.
SecPAL is a logic of authorization for access control decisions in distributed systems.
It has a clear and readable syntax, as well as rich mechanisms for delegation and constraints.
SecPAL has already been used as a basis for other policy languages in areas such as privacy preferences~\cite{Becker:2009ula} and data-sharing~\cite{Aziz:2011vt}.
We present AppPAL as a new version of SecPAL, targeting apps on mobile devices.

\begin{figure}
  \begin{minipage}[t]{0.5\textwidth}
    \begin{lstlisting}[numbers=left,
                       escapeinside={@}{@}]
"alice" says "emma" can-say inf
  App isRunnable. @\label{lst:corporate_1}@

@@

"emma" says "google-play" can-say 0
  App isReputable. @\label{lst:corporate_4}@
    \end{lstlisting}
  \end{minipage}\begin{minipage}[t]{0.5\textwidth}
    \begin{lstlisting}[numbers=left,
                       escapeinside={@}{@},
                       firstnumber=15]
"emma" says "anti-virus-policy" isMetBy(App) @\label{lst:corporate_5}@
  if App isAnApp
  where
    mcAfeeVirusCheck(App) = false.

@@

In \autoref{lst:corporate_1} Alice gives Emma the ability to specify whether an \code{App} (a variable) \code{isRunnable} (a predicate).
She allows her to delegate the decision further if she chooses (\code{can-say inf}).
Next in \autoref{lst:corporate_2} Emma specifies her concerns as policies to be met (the \code{isMetBy()} predicate that takes an app as its argument).
If Emma can be convinced all these policies are met then he will say the \code{App isRunnable}.
In \autoref{lst:corporate_3} and \autoref{lst:corporate_4} Emma specifies that an app meets the \code{reputable-policy} if the \code{App isReputable};
  with \code{"google-play"} specified as the decider of what is buyable or not.
This time Google is not allowed to delegate the decision further (\code{can-say  0}).
In other words Google is not allowed to specify Amazon as a supplier of apps as well.
Google must say what is buyable directly for Emma to accept it.
Emma specifies the \code{"anti-virus-policy"} in \autoref{lst:corporate_5}.
Here we use a constraint.
When checking the policy the \code{mcAfeeVirusCheck} should be run on the \code{App}.
Only if this returns \code{false} will the policy be met.
To specify the \code{"no-tracking-policy"} Emma says that the \code{"no-location-permissions"} rules implement the \code{"no-tracking-policy"} (\autoref{lst:corporate_6}).
Emma specifies this in \autoref{lst:corporate_7} by checking the app is missing two permissions.

Alice wants to install a new app (\code{com.facebook.katana}) on her phone.
To meet Emma's policy the AppPAL policy checker needs to collect statements to show the app meets the \code{isRunnable} predicate.
Specifically it needs:
\begin{itemize}
  \item\code{"emma" says "com.facebook.katana" isAnApp.}
    A simple typing statement that can be generated for all apps as they are encountered.
    This helps keep the number of  assertions in the policy low aiding readability.
  \item\code{"google-play" says "com.facebook.katana" isReputable.}
    Required to convince Emma the app came from a reputable source.
    It should be able to obtain this statement from the Play store as the app is available there.
  \item\code{"emma" says "anti-virus-policy" isMetBy("com.facebook.katana").}
    She can obtain this by running the \ac{av} program on her app.
  \item\code{"emma" says "no-locations-permissions" isMetBy("com.facebook.katana").}
    Needed to show the App meets Emma's no-tracking-policy.
    Emma will say this if after examining the app the location permissions are missing.
\end{itemize}
These last two statements require the checker to do some extra checks to satisfy the constraints.
To get the third statement it must run the \ac{av} program on her app and check the result.
The results from the \ac{av} program may change with time as it's signatures are updated;
  so the checker must re-run this check every time it wants to obtain the statement connected to the constraint.
For the forth statement the checker needs to check the permissions of the app.
It could do this by looking in the \texttt{MANIFEST.xml} inside the app itself, or through the Android package manager if running on device.

In this scenario we have imagined Alice wanting to check the apps as she installs them.
Alternatively we could imagine Emma wanting a personalised app store where all apps sold meet her policy.
With AppPAL this can be implemented by taking an existing store and selectively offering only the apps which will meet the user's policy.
This gives us a \emph{filtered store}.
From an existing set of apps we produce a personalised store that meets a pre-defined policy.

@@

AppPAL is an instantiation of SecPAL for Android app installation policies.
It is implemented as a library for Android and Java.
The parser is implemented using ANTLR4.
Code and build instructions are available from Github\footnote{\url{https://github.com/bogwonch/libAppPAL}}.

The structure of an AppPAL assertion can be seen in \autoref{fig:assertion}.

\begin{figure}
  \centering
  \begin{equation*}\small
    \begin{array}{r l}
      \overbrace{\text{\ttfamily "user"}}^{\text{speaker}} &
      \text{\ttfamily says }\overbrace{\overbrace{\text{\ttfamily App }}^{\text{subject}}\overbrace{\text{\ttfamily isRunnable}}^{\text{predicate}}}^{\text{fact}} \\
      & \overbrace{\text{\ttfamily if App isFree}}^{\text{condition}} \\
      & \overbrace{\text{\ttfamily where hasPermission(App, "INTERNET") = true}}^{\text{constraint}}.
    \end{array}
  \end{equation*}
  \caption{Structure of an AppPAL assertion.}
\label{fig:assertion}
\end{figure}

In the Becker~\etal's paper~\cite{Becker:2006vh} they leave the choice of
predicates, and constraints for their SecPAL open.  With AppPAL we make
explicit our predicates and how they relate to Android.  Some of these
predicates require arguments.  AppPAL policies typically make use of the
predicates and constraints in \autoref{tab:predicates}.  Additional predicates
can be created in the policy files, but adding or modifying constraints
requires a code change\footnote{But a small one.  The Android version of the
\texttt{hasPermission} constraint, for example, uses the Android package
manager to determine what permissions an app requests, but the Java version
uses the Android platform tools.  This required a code change to one function.}

\begin{table}
\begin{tabulary}{\linewidth}{l L}
  \toprule
  Name & Description \\
  \midrule
  App \texttt{isRunnable} & Says an app can be run. \\
  App \texttt{isInstallable} & Says an app can be installed. \\
  App \texttt{isAnApp} & Tells AppPAL that an app exists. \\
  Policy \texttt{isMetBy(}App\texttt{)} & Says a specific policy is met by an app.  This is used to split policies into smaller components which can be reused and composed. \\
  \texttt{hasPermission(}App, Permission\texttt{)} & Constraint for testing whether an app has been granted a permission. \\
  \texttt{BeforeHourOfDay(}time\texttt{)} & Constraint used to test whether we're before an hour in the day. \\
  \bottomrule
\end{tabulary}
\caption{Typical AppPAL predicates and constraints}
\label{tab:predicates}
\end{table}

@@

Splitting the decision about whether an app is runnable into a series of policies that must be met gives us flexibility in how the decision is made.
It allows us to describe multiple means of making the same decision, and provide backup routes when one means fails.
Some static analysis tools are not quick to run.
Even taking minutes to run a battery draining analysis can be undesirable.
If a user wants to download an app quickly they may not be willing to wait to check that a policy is met.

Earlier we described a \emph{no-tracking-policy} to prevent a user's location being tracked.
In Bob's policy we checked this by checking the permissions of the app.
If the app couldn't get access to the GPS sensors (using the permissions) then it met this policy.
Some apps may want to access this data, but may not leak it.
We could use a taint analysis tool to detect this (e.g.~Taintdroid~\cite{Fritz:2013vi}).
Our policy now becomes:

\begin{lstlisting}
"bob" says "no-locations-permissions"
  can-act-as "no-tracking-policy".

"bob" says "no-locations-permissions" isMetBy(App)
  if App isAnApp
  where
    hasPermission(App, "ACCESS_FINE_LOCATION") = false,
    hasPermission(App, "ACCESS_COARSE_LOCATION") = false.

"bob" says "location-taint-analysis"
  can-act-as "no-tracking-policy".

"bob" says "location-taint-analysis" isMetBy(App)
  if App isAnApp
  where
    taintDroidCheckLeak(App, "Location", "Internet") = false.
\end{lstlisting}

Sometimes we might want to use location data.
For instance Bob might want to check that Alice is at her office.
Bob might track Alice using a location tracking app.
Provided the app only talks to Bob, and it uses SSL correctly (which Mallodroid can check for~\cite{Fahl:2012dj}) he is happy to relax the policy.

\begin{lstlisting}
"bob" says "relaxed-no-tracking-policy" canActAs "no-tracking-policy".
"bob" says "relaxed-no-tracking-policy" isMetBy(App)
  if App hasCategory("tracking")
  where
    mallodroidSSLCheck(App) = false,
    connections(App) = "[https://bob.com]".
\end{lstlisting}

This gives us four different ways of satisfying the \emph{no-tracking-policy}:
  with permissions,
  with taint analysis,
  with a relaxed version of the policy,
  or by Bob directly saying the app meets it.
When we come to check the policy if any of these ways give us a positive result we can can stop our search.

@@

Since AppPAL is an instantiation of SecPAL the policy checking rules are the same.
We do not use Becker~\etal's DatalogC~\cite{Li:2003ix}~based translation and evaluation algorithm, as no DatalogC library exists for Android.
We have implemented the rules directly in Java.
Pseudo-code is given in \autoref{fig:pseudocode}.
Like Becker~\etal we make use of an assertion context to store known statements.
We also store intermediate results to avoid re-computation.
On a mobile device memory is at a premium.
We would like to keep the context as small as possible.
For some assertions (like \code{isAnApp}) we derive them by checking the arguments at evaluation time.

% \begin{figure}
% \[\footnotesize
% % AC, D ⊧ A says B can-act-as C   AC, D ⊧ A says C verbphrase
% % ----------------------------------------------------------- (can-act-as)
% %                 AC, D ⊧ A says B verbphrase
% \infer[\text{\tiny can-act-as}]{
%   AC, D \models \text{\tt A says B verbphrase}}{%
%   AC, D \models \text{\tt A says B can-act-as C} & %
%   AC, D \models \text{\tt A says C verbphrase}}
% \]
% \[\footnotesize
% % AC, ∞ ⊧ A says B can-say D fact   AC, D ⊧ B says fact
% % ----------------------------------------------------- (can-say)
% %                   AC, ∞ ⊧ A says fact
% \infer[\text{\tiny can-say}]{
%   AC, \infty \models \text{\tt A says fact}}{%
%   AC, \infty \models \text{\tt A say B can-say $D$ fact} & %
%   AC, D \models \text{\tt B says fact}}
% \]
% \[\footnotesize
% % (A says fact if fact₁, …, factₖ, c) ∈ AC
% % AC, D ⊧ A says factᵢθ  ∀i ∈ {1…k}    ⊧cθ    vars(factθ) = ∅
% % ----------------------------------------------------------- (cond)
% %                    AC, D ⊧ A says factθ
% \infer[\text{\tiny cond}]{
%   AC, D \models \text{\tt A says fact$\theta$}}{
%   \begin{array}{c}
%     (\text{\tt A says fact if fact$_1$, \ldots, fact$_k$ where c}) \in AC \\
%     \forall i \in 1\cdots k.\;AC, D \models \text{\tt A says fact$_i\theta$}
%   \end{array} &
%   \models \mathtt{c}\theta &
%   \text{\sl vars}(\text{\tt fact$\theta$}) = \emptyset}
% \]
% \caption{AppPAL's evaluation rules.}
% \label{fig:rules}
% \end{figure}

This gives us greater control of the evaluation and how the assertion context is created.
For example, when checking the \code{isAnApp} predicate;
  we can fetch the assertion that the subject is an app based on the app in question.
Similarly when we use a statement from \emph{Emma} that \emph{Google-Play can-say} whether an app is buyable;
  it is sensible to go fetch from the store whether the app is saleable and make Google say it then and there.

@@

\section{Conclusions and further work \comment{$\frac{1}{2}$ page}}

We have presented AppPAL: an authorization logic for describing app installation policies.
The language is implemented in Java and runs on Android using a custom evaluation algorithm.
This lets us to enforce app installation policies on Android devices.
We have shown how the language can be used to describe an app installation policy;
  and given brief descriptions of how other policies might be described.

>>>>

\begin{abstract}
  Users must judge apps by the information shown to them by the store.
  Employers rely on employees enforcing company policies correctly on devices at their workplace.
  Some users take time to pick apps with care whilst others do not.
  Users feel frustrated when they realise what data the apps have access to.
  They want greater control over what data they give away but they do not want to spend time reading permissions lists.
  We present AppPAL: a policy language to enforce policies about apps.
  AppPAL policies use statements from third parties and delegation relationships to give us a rigorous and flexible framework for enforcing and comparing app policies; and the trust relationships surrounding them.
\end{abstract}

@@

Finding the right apps can be tricky.
Users need to discover which apps are well written, which are not going to abuse their data
  and to find the apps which suit how they want to use their device.
This can be difficult as it isn't obvious how apps use the data each has access to.

App stores give some information about their apps; descriptions of the app and screenshots as well as review scores.
Android apps show a list of permissions when they're first installed.
Soon new apps will display permissions requests when the app first tries to access sensitive data (such as contacts or location information).
Users do not understand how permissions relate to their device~\cite{Felt:2012hm,Thompson:2013eb}.
Ultimately the decision which apps to use and which permissions to grant must be made by the user.

Not all apps are suitable.
Many \ac{pup} are being propagated for Android devices~\cite{Truong:2014bi,Svajcer:2013tp}.
Employees are increasingly using their own phones for work (a system known as \ac{byod}).
An employer may wish to restrict which apps their employees can use.
The IT department may set a policy to prevent information leaks.
Some users worry apps will misuse their personal data;
  such a user avoids apps which can access their location, or address book.
They may apply their own personal security policy when downloading and running apps.

These policies can only be enforced by the users continuously making the correct decision when prompted about apps.
This is error-prone.
We believe this can be improved.
An alternative would be to write the policy down and make the computer enforce it.
To implement this we use a logic of authorization.
The policy is written in the logic and enforced by checking the policy is satisfied.

We present AppPAL, an authorization logic for reasoning about apps.
The language is an instantiation of Becker~\etal's SecPAL~\cite{Becker:2006vh} with constraints and predicates that allow us to decide which apps to run or install.
The language allows us to reason about apps using statements from third parties.
The implementation allows us to enforce the policies on a device.
We can express trust relationships amongst these parties; use constraints to do additional checks.
This lets us enforce deeper and more complex policies than existing tools such as Kirin~\cite{Enck:2009ko}.

@@

"alice" says App isRunnable
  if "work-policy" isMetBy(App)
  where BeforeHourOfDay("17") = true.
\end{lstlisting}
We can delegate policy specification to third parties or roles, and assign principals to those rolls:
\begin{lstlisting}
"alice" says "it-department" can-say 0 "work-policy" isMetBy(App).
"alice" says "alice" can-act-as "it-department".
\end{lstlisting}
We can write policies specifying which permissions an app must or must not have by its app store categorization:
    \begin{lstlisting}
"alice" says App isRunnable
   if "permissions-policy" isMetBy(App).
"alice" says "permissions-policy" isMetBy(App)
   if App isAnApp
   where
      category(App, "Photography"),
      hasPermission(App, "LOCATION") = false,
      hasPermission(App, "CAMERA") = true.
    \end{lstlisting}

@@

\noindent
In this paper we have:
\begin{itemize}
  \item
    Described a scenario where an employer has a policy they want to enforce for their employees~(\autoref{sec:problem});
    and shown how the employer's policy could be implemented using AppPAL~(\autoref{sec:idea}) and installed on Alice's phone.
  \item
    Implemented the AppPAL language as a library on Android and the JVM.
    We show how the language can describe properties of Android apps and Android security policies (\autoref{ssec:language}) and how we can check policies (\autoref{ssec:eval}).
  \item
    Shown the need for policy tools by demonstrating the privacy paradox holds for user privacy policies with app installations (\autoref{sec:demonstation}).
\end{itemize}

@@

To ensure this policy is met Alice promises to follow it.
She might even sign a document promising never to break the rules within the policy.
This is error-prone---what if she makes a mistake or misses an app that breaks her policy?
Emma's policy could be implemented using existing tools.
\emph{Google's Device Policy for Android}\footnote{\url{https://play.google.com/store/apps/details?id=com.google.android.apps.enterprise.dmagent}} could configure Alice's device to disallow apps from outside the Google Play Store and let Emma set the permissions granted to each app~\cite{AndroidMPermission:2015uq}).
AppPAL is designed to build on these tools, but make the trust and delegation relationships explicit and provide a rigorous means for deciding whether an app meets a policy.
Various tools such as AppGuard~\cite{Backes:2012vm}, Dr.~Android~\&~Mr.~Hide~\cite{Jeon:2012ki} or AppFence~\cite{Hornyack:2011wq} can control the permissions or data an app can get.
These could be used to used to ensure no location data is ever obtained.
Alternately other tools like Kirin~\cite{Enck:2009ko}, Flowdroid~\cite{Fritz:2013vi} or DroidSafe~\cite{Gordon:2015et} could check that the locations are never leaked to the web.
Various anti-virus programs are available for Android---one could be installed on Alice's phone checking against McAfee's signatures.

Whilst we could implement Emma's policy using existing tools it is a clumsy solution---they are not flexible.
If Emma changes her policy or Alice changes jobs she needs to recheck her apps and then alter or remove the software on her phone to ensure compliance.
It isn't clear what an app must do to be run, or what checks have been done if it is already running on the phone.
The relationship between Alice (the user), Emma (the policy setter) and the tools Emma trusts to implement her policy isn't immediately apparent.

What happens when Alice goes home?
Emma shouldn't be able to overly control what Alice does in her private life.
Alice might not be allowed to use location tracking apps at work but at home she might want to (to meet friends, track jogging routes or find restaurants for example).
Some mobile OSs, such as iOS and the latest version of Android, allow app permissions to be enabled and disable at run time.
Can we enforce different policies at different times or locations?

Our research looks at the problem of picking software.
Given there are some apps you want to install and run and others you do not:
  how can you express your preferences in such a way that they can be enforced automatically?
How can we translate policy documents from natural language into a machine checkable form?
How can we show the trust relationships used to make these decisions clearly and precisely?

@@

We propose a change to the ecosystem, shown in \autoref{fig:ecosystem}.
People have policies which are enforced by AppPAL on their devices.
They can be composed with policies from employers or others to create enhanced devices that ensure apps meet the policies of their owners.
The device can make use of vetting services which run tools to infer complex properties about apps.
Users can buy from enhanced stores which ensure the only apps they sell are the apps which meet the explicitly specified store policies.
Developers could decide which stores to sell their apps in on the basis of policies about stores.

@@

In \autoref{sec:problem} Alice and Emma had policies they wanted to enforce but no means to do so.
Instead of using several different tools to enforce Emma's policy disjointedly, we could use an authorization logic.
In \autoref{lst:corporate} we give an AppPAL policy implementing Emma's app concerns on Alice's phone.
% A pictorial representation, is given in \autoref{fig:emmas_policy}; color splits statements by each speaker.

SecPAL is a logic of authorization for access control decisions in distributed systems.
It has a clear and readable syntax, as well as rich mechanisms for delegation and constraints.
SecPAL has already been used as a basis for other policy languages in areas such as privacy preferences~\cite{Becker:2009ula} and data-sharing~\cite{Aziz:2011vt}.
We present AppPAL as a modified form of SecPAL, targeting apps on mobile devices.

\begin{figure}
  \begin{minipage}[t]{0.5\textwidth}
    \begin{lstlisting}[numbers=left,
                       escapeinside={@}{@},
                       basicstyle=\scriptsize\ttfamily,
                       stringstyle=\scriptsize\sffamily,
                       keywordstyle=\scriptsize\slshape]
"alice" says "emma" can-say inf
  App isRunnable. @\label{lst:corporate_1}@

@@

"emma" says "google-play" can-say 0
  App isReputable. @\label{lst:corporate_4}@
    \end{lstlisting}
  \end{minipage}\begin{minipage}[t]{0.5\textwidth}
    \begin{lstlisting}[numbers=left,
                       escapeinside={@}{@},
                       firstnumber=15,
                       basicstyle=\scriptsize\ttfamily,
                       stringstyle=\scriptsize\sffamily,
                       keywordstyle=\scriptsize\slshape]
"emma" says "anti-virus-policy" isMetBy(App) @\label{lst:corporate_5}@
  if App isAnApp
  where
    mcAfeeVirusCheck(App) = false.

@@

In \autoref{lst:corporate_1} Alice gives Emma the ability to specify whether an \code{App} (a variable) \code{isRunnable} (a predicate).
She allows her to delegate the decision further if she chooses (\code{can-say inf}).
Next in \autoref{lst:corporate_2} Emma specifies her concerns as policies to be met (the \code{isMetBy()} predicate that takes an app as its argument).
If Emma can be convinced all these policies are met then she will say the \code{App isRunnable}.
In \autoref{lst:corporate_3} and \autoref{lst:corporate_4} Emma specifies that an app meets the \code{reputable-policy} if the \code{App isReputable};
  with \code{"google-play"} specified as the decider of what is buyable or not.
This time Google is not allowed to delegate the decision further (\code{can-say 0}).
In other words Google is not allowed to specify Amazon as a supplier of apps as well.
Google must say what is buyable directly for Emma to accept it.
Emma specifies the \code{"anti-virus-policy"} in \autoref{lst:corporate_5}.
Here we use a constraint.
When checking the policy the \code{mcAfeeVirusCheck} should be run on the \code{App}.
Only if this returns \code{false} will the policy be met.
To specify the \code{"no-tracking-policy"} Emma says that the \code{"no-location-permissions"} rules implement the \code{"no-tracking-policy"} (\autoref{lst:corporate_6}).
Emma specifies this in \autoref{lst:corporate_7} by checking the app is missing two permissions.

Alice wants to install a new app (\code{com.facebook.katana}) on her phone.
To meet Emma's policy the AppPAL checker needs to collect statements to show the app meets the \code{isRunnable} predicate.
Specifically it needs:
\begin{itemize}
  \item\code{"emma" says "com.facebook.katana" isAnApp.}
    A simple typing statement that can be generated for all apps as they are encountered.
    This helps keep the number of assertions in the policy low aiding readability.
  \item\code{"google-play" says "com.facebook.katana" isReputable.}
    Required to convince Emma the app came from a reputable source.
    It should be able to obtain this statement from the Play store as the app is available there.
  \item\code{"emma" says "anti-virus-policy" isMetBy("com.facebook.katana").}
    She can obtain this by running the \ac{av} program on her app.
  \item\code{"emma" says "no-locations-permissions" isMetBy("com.facebook.katana").}
    Needed to show the App meets Emma's no-tracking-policy.
    Emma will say this if after examining the app the location permissions are missing.
\end{itemize}
These last two statements require the checker to do some extra checks to satisfy the constraints.
To get the third statement it must run the \ac{av} program on her app and check the result.
The results from the \ac{av} program may change with time as it's signatures are updated;
  so the checker must re-run this check every time it wants to obtain the statement connected to the constraint.
For the forth statement the checker needs to check the permissions of the app.
It could do this by looking in the \texttt{MANIFEST.xml} inside the app itself, or through the Android package manager if it is running on a device.

In this scenario we have imagined Alice wanting to check the apps as she installs them.
We could also imagine Emma wanting a personalised app store where all apps sold meet her policy.
With AppPAL this can be implemented by taking an existing store and selectively offering only the apps which will meet the user's policy.
This gives us a \emph{filtered store} which, from an existing set of apps, we get a personalised store that only sells apps that meet a policy.

@@

AppPAL is implemented as a library for Android and Java.
The parser is implemented using ANTLR4.
The structure of an AppPAL assertion can be seen in \autoref{fig:assertion}.

\begin{figure}
  \newcommand{\bracetext}[1]{\text{\sffamily #1}}
  \newcommand{\smalltext}[1]{\text{\ttfamily\footnotesize #1}}
  \centering
  \begin{equation*}\small
    \begin{array}{r l}\footnotesize
      \overbrace{\smalltext{"user"}}^{\bracetext{speaker}} &
      \smalltext{ says }\overbrace{\overbrace{\smalltext{ App }}^{\bracetext{subject}}\overbrace{\smalltext{ isRunnable}}^{\bracetext{predicate}}}^{\bracetext{fact}} \\
      & \overbrace{\smalltext{ if App isFree}}^{\bracetext{condition}} \\
      & \overbrace{\smalltext{ where hasPermission(App, "INTERNET") = true}}^{\bracetext{constraint}}.
    \end{array}
  \end{equation*}
  \caption{Structure of an AppPAL assertion.}
\label{fig:assertion}
\end{figure}

In Becker~\etal's work~\cite{Becker:2006vh} they leave the choice of predicates, and constraints for their SecPAL open.
With AppPAL we make explicit our predicates and how they relate to Android.
AppPAL policies can make use of the predicates and constraints in \autoref{tab:predicates}.
Additional predicates can be created in the policy files, but adding or modifying constraints required a code change\footnote{The Android version of the \texttt{hasPermission} constraint, for example, uses the Android package manager to determine what permissions an app requests, but the Java version uses the Android platform tools.  This requires a small code change.}

\begin{table}
\begin{tabulary}{\linewidth}{l L}
  \toprule
  Name & Description \\
  \midrule
  App \texttt{isRunnable} & Says an app can be run. \\
  App \texttt{isInstallable} & Says an app can be installed. \\
  App \texttt{isAnApp} & Tells AppPAL that an app exists. \\
  Policy \texttt{isMetBy(}App\texttt{)} & Says a specific policy is met by an app.  This is used to split policies into smaller components which can be reused and composed. \\
  \texttt{hasPermission(}App, Permission\texttt{)} & Constraint for testing whether an app has been granted a permission. \\
  \texttt{BeforeHourOfDay(}time\texttt{)} & Constraint used to test whether we're before an hour in the day. \\
  Tool\texttt{Check(}App, Property) &  General form of a constraint used to run a static analysis tool over an app. \\
  \bottomrule
\end{tabulary}
\caption{Typical AppPAL predicates and constraints}
\label{tab:predicates}
\end{table}

@@

Splitting the decision about whether an app is runnable into a series of policies that must be met gives us flexibility in how the decision is made.
It allows us to describe multiple means of making the same decision, and provide backup routes when one fails.
Some static analysis tools are not quick to run.
Even taking minutes to run a battery draining analysis can be undesirable.
If a user wants to download an app quickly they may not be willing to wait to check that a policy is met.

In \autoref{sec:problem} and \autoref{sec:idea} we described a \emph{no-tracking-policy} to prevent a user's location being leaked.
In Emma's policy we checked this using the app's permissions.
If the app couldn't get access to the GPS sensors (using the permissions) then it met this policy.
Some apps may want to access this data, but may not leak it.
We could use a taint analysis tool to detect this (e.g.~Taintdroid~\cite{Fritz:2013vi}).
Our policy now becomes:

\begin{lstlisting}
"emma" says "no-locations-permissions"
  can-act-as "no-tracking-policy".

"emma" says "no-locations-permissions" isMetBy(App)
  if App isAnApp
  where
    hasPermission(App, "ACCESS_FINE_LOCATION") = false,
    hasPermission(App, "ACCESS_COARSE_LOCATION") = false.

"emma" says "location-taint-analysis"
  can-act-as "no-tracking-policy".

"emma" says "location-taint-analysis" isMetBy(App)
  if App isAnApp
  where
    taintDroidCheck(App, "Location", "Internet") = false.
\end{lstlisting}

Sometimes we might want to use location data.
For instance Emma might want to check that Alice is at her office.
Emma might track Alice using a location tracking app.
Provided the app only talks to Emma, and it uses SSL correctly (which Mallodroid can check for~\cite{Fahl:2012dj}) she is happy to relax the policy.

\begin{lstlisting}
"emma" says "relaxed-no-tracking-policy" canActAs "no-tracking-policy".
"emma" says "relaxed-no-tracking-policy" isMetBy(App)
  if App hasCategory("tracking")
  where
    mallodroidSSLCheck(App) = false,
    connectionsCheck(App, "[https://emma.com]") = true.
\end{lstlisting}

This gives us four different ways of satisfying the \emph{no-tracking-policy}:
  with permissions,
  with taint analysis,
  with a relaxed version of the policy,
  or by Emma directly saying the app meets it.
When we come to check the policy if any of these ways give us a positive result we can stop our search.

@@

AppPAL has the same policy inference rules as SecPAL.
We do not use Becker~\etal's DatalogC~\cite{Li:2003ix}~based checking algorithm, as no DatalogC library exists for Android.
We have implemented the rules directly in Java.
Pseudo-code is given in \autoref{fig:pseudocode}.

Like Becker~\etal~we make use of an assertion context to store known statements.
We also cache intermediate results, and ground facts to avoid re-computation.
On a mobile device memory is at a premium.
We would like to keep the context as small as possible.
For some assertions (like \code{isAnApp}) we derive them by checking the arguments at evaluation time.
This gives us greater control of the evaluation and how the assertion context is created.
For example, when checking the \code{isAnApp} predicate;
  we can fetch the assertion that the subject is an app based on the app in question.
Similarly when we use a statement from \emph{Emma} that \emph{Google-Play can-say} whether an app is buyable;
  it is sensible to go fetch from the store whether the app is saleable and make Google say it then and there.

@@

\section{Conclusions and further work}

We have presented AppPAL: an authorization logic for describing app installation policies.
AppPAL lets us enforce app installation policies on Android devices.
We have shown how the language can be used to describe an app installation policy;
  and given brief descriptions of how other policies might be described.
